Bayesian non-parametric inference for multivariate peaks-over-threshold models

### 2023-05-17 pointbypoint.tex

We note in passing that the moment conditions for the spectral distribution, say 
    $\Phi$, in the two-dimensional case, using the infinity norm, imply that 
    \[
        \int_0^{\pi/4} \Phi(d\theta) = \int_{\pi/4}^{\pi/2}
    \Phi(d\theta).
    \]
    The former looks like a fairly strong symmetry restriction that is unlike to 
    be realistic in practice. Thus, imposing moment conditions would probably 
    result in a very restrictive model.

\item \emph{Main contribution}: While I appreciate the interesting construction 
    based on the DPM of projected gammas (which itself could be of use for other 
    contexts beyond extremes?), one wonders about what would be the main added 
    value of it with respect to other existing Bayesian approaches 
    (\cite{boldi2007},\cite{guillotte2011},\cite{SaNa2014},\cite{hanson2017}).
    Keeping in mind the computational scope of Statistics and Computing, one 
    wonders if the main advantage is, for example, a computational one? This 
    needs to be clarified—especially in light of the previous comment on the 
    moment constraint. Indeed, one could use any of the previous approaches with 
    pseudo-angles based on the infinity norm.

We have clarified the distinction between our approach and the prior contributions
    mentioned.  Setting aside \cite{guillotte2011} which is a strictly bivariate
    model, the named approaches consider batch extrema and by construction must
    respect the moment constraint.  As noted, our approach is able to set that 
    aside, allowing consideration of a wider class of distributions of pseudo-angles.

@@

    % \comment{Hanson is clearly different.  They're using Bernstein polynomials 
    % effectively defined on $\mathbb{S}_1^{d-1}$.  We can demonstrate that 
    % projected gammas built on $\mathbb{S}_{p}$, $p \geq 10$ and $\mathbb{S}_1$ 
    % create \emph{very} different gamma shape parameters, for data created by 
    % the \emph{same} distribution.  I need to check again the other references, 
    % but based on their follow-on comment, I'm guessing the previous point needs 
    % to be addressed before we can properly justify this point.}

    % \bruno{Based on our conversation, you need to look carefully into each one 
    % of these references. it seems that most of them do not develop a PoT 
    % approach. But also, we use a kernel whose natural support is the positive 
    % cone (without using transformations) to build a mixture WRT to a random
    % measure. This provides the flexibility of mixture for a methods that has 
    % computational advantages and scales to moderately large problems, as 
    % illustrated in the paper.}

    % \comment{\cite{Hanson2017}: Bernstein polynomials on $\mathbb{S}_1^{d-1}$, 
    % modelling multivariate Frechets.  \cite{boldi2007}: Marginal Frechets, also 
    % operating on $\mathbb{S}_1^{d-1}$.  \cite{guillotte2011}: A strictly bivariate model.  
    % \cite{SaNa2014}: }

@@

\item \emph{Numerical accuracy}:  Keeping in mind the computational scope of 
    Statistics and Computing one would hope for more numerical work than what is 
    presented in §5.1. As far as I understand what is presented are single run 
    experiments for each $d$? Monte Carlo evidence on the performance of the 
    procedure would be appreciated. On top of this, the current numerical exercise
    ignores the uncertainty that stems from step 1 of the procedure 
    (i.e. estimating the margins).

To address this concern, we have revised the simulation study to feature the 
    average rise in energy score over baseline, averaged over 10 replications.  
    The simulated datasets are generated as mixtures of multivariate lognormals, 
    projected onto the unit sphere.  As you say, this paper focuses on estimating
    the spectral distribution, and ignores uncertainty from estimating marginal
    parameters.  We have clarified as such.

### 2023-05-23 results.tex
Updates from Overleaf

\begin{figure*}[htb]
    \centering
    \caption{Energy score (top row) and Posterior Predictive Loss (bottom row) for various models on simulated data with ascending count of mixture components (indicated by plot heading) and number of dimensions (indicated by horizontal axis).  Gen indicates baseline.\label{fig:simpples}}
    \includegraphics[height=3in, width = \textwidth]{./images/simulation_score}
\end{figure*}

@@

\subsection{Simulations\label{subsec:simulated}}

To evaluate our proposed approach for angular measure estimation 
    we consider simulated datasets on $\mathbb{S}_{\infty}^{d-1}$, for 
    $d$ ranging between 3 and 20. We generated each dataset from a mixture of projected Gammas, with 
    the number of mixture components ranging between 3 and 12.  The generation procedure is detailed 
    in Algorithm~\ref{algo:simulated}.
    \begin{algorithm}[ht]  % algorithm using algorithm package
        \caption{Simulated Angular Dataset Generation Routine\label{algo:simulated}}
        \begin{algorithmic}
        \For{$n_{\text{mix}}$ in $[3, 6, 9, 12]$}
            \State Generate $n_{\text{mix}} \times 20$ shape Parameters $\alpha$
            \State Generate $n_{\text{mix}} \times 20$ Rate Parameters $\beta$
            \State Generate $500$ Mixture Component Identifiers $\delta$
            \For{$i$ in $1,\ldots,500$}
                \State Generate ${\bf X}_i \sim \prod_{\ell = 1}^d\text{Ga}\left(X_{il}\mid\alpha_{\delta_i,l},\beta_{\delta_i, l}\right)$
            \EndFor
            \For{$n_{\text{col}}$ in $[3,6,12,20]$}
                \State Project columns 1 to $n_{\text{col}}$ of ${\bf X}$ onto $\mathcal{S}_{\infty}^{n_{\text{col}} - 1}$ and save.
            \EndFor
        \EndFor
        \end{algorithmic}
    \end{algorithm}

Each one of the shape parameters $\alpha$ were generated 
$\alpha = \alpha_0 + \alpha_1$ where
  $\alpha_0 \sim \text{Unif}(\alpha_0\mid 0,4)$, $\alpha_1\sim \text{Gamma}(\alpha_1\mid 1,1)$.
The scale parameters $\beta$ were obtained as $\beta\sim\text{Unif}(\beta\mid 0.25, 2.5)$.

Figure~\ref{fig:simpples}, shows two sets of panels that illustrate the comparison
between four different models used to fit the simulated data. We produced 16 
different data sets, one for each of four different numbers of mixture components 
and four different dimensions. We considered three different models: DP mixtures 
of projected gamma; Projected restricted gamma; and projected restricted gamma 
with a multivariate log-normal prior. We calculated $S^\infty_{\text{PPL}}$ and 
$S_{\text{ES}}$ using predictive samples from each of fitted models. To provide 
a comparative baseline, we calculated the scores using samples generated 
directly using the same model and parameters that produced the simulated 
observations. The results are denoted as \emph{Gen} in the figure. 
We observe that the projected
  restricted Gamma model dominates in most situations, but the difference between projected restricted
  Gamma and the other gamma-based models tends to shrink as the number of dimensions increases.
  Alternatively, that difference appears to grow as the number of mixture components increases. 
  The results from both scoring criteria are comparable.

>>>>

\begin{figure*}[htb]
    \centering
    % \caption{Energy score (top row) and Posterior Predictive Loss (bottom row) for various models on simulated data with ascending count of mixture components (indicated by plot heading) and number of dimensions (indicated by horizontal axis).  Gen indicates baseline.\label{fig:simpples}}
    \caption{Average energy score rise over baseline (on $\mathbb{S}_{\infty}^{d-1}$) for various models on simulated data, with ascending count of mixture components (indicated by plot heading) and number of dimensions (indicated by horizontal axis).  Note that pairwise betas is a moment-restricted model.\label{fig:simpples}}
    \includegraphics[height=3in, width = \textwidth]{./images/sim_es_rise}
\end{figure*}

@@

\subsection{Simulations\label{subsec:simulated}}
To evaluate our proposed approach for angular measure estimation 
    we consider simulated datasets on $\mathbb{S}_{\infty}^{d-1}$,
    for values of $d$ ranging between 2 and 32.  We generated each
    dataset as a mixture of multivariate log-normal distributions,
    projected onto $\mathbb{S}_p^{d-1}$. The generation procedure is detailed 
    in Algorithm~\ref{algo:simulated}.

\begin{algorithm}[ht]  % algorithm using algorithm package
        \caption{Simulated Angular Dataset Generation Routine\label{algo:simulated}}
        \begin{algorithmic}
        \For{$n_{\text{iter}}$ in $[1,\ldots,10]$}
        \For{$n_{\text{mix}}$ in $[1, 2, 4, 8]$}
            \For{$j$ in $1,\ldots,n_{\text{mix}}$}
                \State Generate $\mu_{j} \sim \mathcal{N}\left(\bm{0},\bm{I}\right)$
                \State Generate $\Sigma_{j}\sim\mathcal{IW}\left(70,70 \bm{I}\right)$
            \EndFor
            \State Generate $\pi\sim\text{Dirichlet}(\bm{10})$
            \For{$i$ in $1,\ldots,1000$}
                \State Generate $\delta_i \sim \text{Categorical}(\pi)$
                \State Generate $\bm{X}_i \sim \mathcal{LN}\left(\mu_{[\delta_i]},\Sigma_{[\delta_i]}\right)$
            \EndFor
            \For{$n_{\text{col}}$ in $[2,4,8,16,24,32]$}
                \State Project columns 1 to $n_{\text{col}}$ of ${\bf X}$ onto $\mathcal{S}_{\infty}^{n_{\text{col}} - 1}$ and save.
            \EndFor
        \EndFor
        \EndFor
        \end{algorithmic}
    \end{algorithm}

where $\mu_j$, $\Sigma_j$ the parameters of the log-normal mixture component 
    distribution; $\pi$ the probability vector assigning weight to each mixture 
    component, and $\delta_i$ the mixture component identifier for each simulated 
    observation.

Figure~\ref{fig:simpples}, shows the average rise over baseline in energy score as
    calculated on $\mathbb{S}_{\infty}^{d-1}$ using the kernel metric described
    in Proposition~\ref{prop:rot}, for models trained on simulated data.  After training
    a model, a posterior predictive dataset is generated, and the energy score is
    calculated as a Monte Carlo approximation of Equation~\ref{eq:es}.
    The \emph{baseline} value is the energy score of a new dataset from the same
    generating distribution as the training dataset, evaluated against the training 
    dataset.  We considered a total of four variations of the Dirichlet-process mixture
    of projected gammas model: whether to restrict $\beta_{j\ell} := 1$ as the projected
    \emph{restricted} gamma (DP-PRG) akin to a projected Dirichlet, or allow them to be 
    learned with a gamma prior (DP-PG); and for the shape parameters whether to use a 
    gamma centering distribution (-G) or a log-normal centering distribution (-LN).
    For comparison we also included two other distributions on $\mathbb{S}_{1}^{d-1}$:
    The Dirichlet distribution (with gamma priors), and the pairwise betas distribution
    developed in~\cite{COOLEY2010}.  Note that the pairwise betas model respects
    the moment restriction generally assumed.\makenote{This phrasing needs to be changed.
    The point I am trying to make is that because the moment-restriction assumption is 
    discarded, a moment-restricted model is inadequate, as is borne out in the
    simulation study.}  We observe that the projected gamma models dominate, and within
    that class of model, the choice of centering distribution is irrelevant.  For the
    projected \emph{restricted} gamma models, the choice of log-normal centering
    distribution induces

shows two sets of panels that illustrate the comparison
between four different models used to fit the simulated data. We produced 16 
different data sets, one for each of four different numbers of mixture components 
and four different dimensions. We considered three different models: DP mixtures 
of projected gamma; Projected restricted gamma; and projected restricted gamma 
with a multivariate log-normal prior. We calculated $S^\infty_{\text{PPL}}$ and 
$S_{\text{ES}}$ using predictive samples from each of fitted models. To provide 
a comparative baseline, we calculated the scores using samples generated 
directly using the same model and parameters that produced the simulated 
observations. The results are denoted as \emph{Gen} in the figure. 
We observe that the projected
  restricted Gamma model dominates in most situations, but the difference between projected restricted
  Gamma and the other gamma-based models tends to shrink as the number of dimensions increases.
  Alternatively, that difference appears to grow as the number of mixture components increases. 
  The results from both scoring criteria are comparable.
